Comparative vaccine effectiveness (ChAdOx1 versus BNT162b2) in Health and Social Care workers

### 2021-08-10 analysis/report/effectiveness_report_vaccinatedtime.Rmd
add vaxtime-only report

fs::dir_create(output_dir, "report", "figures")

knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(fs::path(output_dir, "report", "figures"), "/")
)

@@

Time-dependent Cox models are used to estimate the relative hazard of each outcome for ChAdOx1 versus BNT162b2 vaccine recipients.
Modelled covariates include age, sex, ethnicity, deprivation, evidence of prior covid infection, asthma, clinically “at risk”, number of recorded comorbidities and severe mental illness.
NUTS1 regions are used as a stratification variable to account for geographical variation in event rates for the outcomes of interest (primarily driven by changes in local infection rates).

Comparative effectiveness is estimated at weekly post-vaccination intervals to account for the potential differences in vaccine protection over time between the two brands.

Two time-scales are used for triangulation:

* On the *calendar time-scale*, time zero is `r format(as.Date(gbl_vars$start_date_az), "%d %B %Y")`, and people only contribute follow-up time once they have been vaccinated (i.e., delayed study entry).

* On the *treatment time-scale*, time zero is the date of vaccination. Calendar-time is included as an additional covariate using restricted cubic spline with knots at the first and second tertile, plus boundary knots.

@@

```{r, curves, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', results='asis'}

@@

  calendar_effects <- read_rds(here("output", "models", outcome, "calendar", glue("report{modeltype}_effects{suffix}.rds"))) %>%
    mutate(timescale = "Calendar time")
  timesincevax_effects <- read_rds(here("output", "models", outcome, "timesincevax", glue("report{modeltype}_effects{suffix}.rds"))) %>%
    mutate(timescale = "Vaccinated time")

  tidy_effects <- bind_rows(calendar_effects, timesincevax_effects)

  plot_effects <-
    ggplot(data = tidy_effects) +
    geom_point(aes(y=exp(estimate), x=term_midpoint, colour=model_name), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=exp(conf.low), ymax=exp(conf.high), x=term_midpoint, colour=model_name), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(cols=vars(timescale))+
    scale_x_continuous(breaks=unique(tidy_effects$term_left), limits=c(min(tidy_effects$term_left), max(tidy_effects$term_right)+1), expand = c(0, 0))+
    scale_y_log10(
    breaks=c(0.1, 0.125, 0.25, 0.33, 0.5, 0.75, 1, 1.33, 2, 3, 4, 8,10),
    sec.axis = dup_axis(name="favours\n <- BNT162b2 / ChAdOx1 ->", breaks = NULL)
  )+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=2))+
    labs(
      y="Hazard ratio",#\nChAdOx1 / BNT162b2",
      x="Days since vaccination",
      colour=NULL
     ) +
    theme_bw()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),

@@

  calendar_curve <- read_rds(here("output", "models", outcome, "calendar", glue("reportplr_cmlinccurves_{spline}.rds"))) %>%
    mutate(timescale = "Calendar time")
  timesincevax_curve <- read_rds(here("output", "models", outcome, "timesincevax", glue("reportplr_cmlinccurves_{spline}.rds"))) %>%
    mutate(timescale = "Vaccinated time")

  curves <- bind_rows(calendar_curve, timesincevax_curve)

  plot_cmlinc <-
    ggplot(curves %>% filter(model=="3"))+
    geom_step(aes(x=tstop, y=1-survival, group=vax1_az_descr, colour=vax1_az_descr))+
    facet_grid(cols=vars(timescale))+
    scale_x_continuous(
      breaks = seq(0,7*52,by=14),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0)
    )+
    scale_colour_brewer(type="qual", palette="Set2")+
    labs(
      x="Days since first dose",
      y="Cumulative risk",
      colour=NULL
    )+
    theme_bw()+
    theme(
      legend.position=c(0.05,.95),
      legend.justification = c(0,1),
      axis.text.x.top=element_text(hjust=0),

@@

>>>>

fs::dir_create(output_dir, "report", "figures-vaxtime")

knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(fs::path(output_dir, "report", "figures-vaxtime"), "/")
)

@@

Time-dependent Cox models are used to estimate the relative hazard of each outcome for ChAdOx1 versus BNT162b2 vaccine recipients.
Modelled covariates include age, sex, ethnicity, deprivation, evidence of prior covid infection, asthma, clinically “at risk”, number of recorded comorbidities and severe mental illness.
NHS regions are used as a stratification variable to account for geographical variation in event rates for the outcomes of interest (primarily driven by changes in local infection rates).

Comparative effectiveness is estimated at weekly post-vaccination intervals to account for the potential differences in vaccine protection over time between the two brands. 
Time zero is the date of vaccination. Calendar-time is included as an additional covariate using restricted cubic spline with knots at the first and second tertile, plus boundary knots.

@@

```{r, curves_vaxtime, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', results='asis'}

@@

  tidy_effects <- read_rds(here("output", "models", outcome, "timesincevax", glue("report{modeltype}_effects{suffix}.rds")))

  plot_effects <-
    ggplot(data = tidy_effects) +
    geom_point(aes(y=exp(estimate), x=term_midpoint, colour=model_name), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=exp(conf.low), ymax=exp(conf.high), x=term_midpoint, colour=model_name), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    scale_x_continuous(breaks=unique(tidy_effects$term_left), limits=c(min(tidy_effects$term_left), max(tidy_effects$term_right)+1), expand = c(0, 0))+
    scale_y_log10(
    breaks=c(0.1, 0.125, 0.25, 0.33, 0.5, 0.75, 1, 1.33, 2, 3, 4, 8,10),
    sec.axis = dup_axis(name="favours\n <- BNT162b2 / ChAdOx1 ->", breaks = NULL)
  )+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=2))+
    labs(
      y="Hazard ratio",#\nChAdOx1 / BNT162b2",
      x="Days since vaccination",
      colour=NULL
     ) +
    theme_bw()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),

@@

  curves <- read_rds(here("output", "models", outcome, "timesincevax", glue("reportplr_cmlinccurves_{spline}.rds")))

  plot_cmlinc <-
    ggplot(curves)+
    geom_step(aes(x=tstop, y=1-survival, group=vax1_az_descr, colour=vax1_az_descr))+
    scale_x_continuous(
      breaks = seq(0,7*52,by=14),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0)
    )+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=2))+
    labs(
      x="Days since first dose",
      y="Cumulative risk",
      colour=NULL
    )+
    theme_bw()+
    theme(
      legend.position=c(0.05,.95),
      legend.justification = c(0,1),
      axis.text.x.top=element_text(hjust=0),

@@

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
